Appendix 4: Model Comparison Code For Manuscript

### 2025-11-10 modelcomparison_code.Rmd
demographics and cor plot and stars

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(worcs)
library(tidyverse)
library(gtsummary)
library(likert)
library(here)
library(corrplot)
library(ggstats)
library(lavaan)
library(lavaanPlot)
library(kableExtra)
library(naniar)

library(gridExtra)
#library(semPlot) # update plots to semPlot later
#library(blavaan) # could not get it to work with this model and synthetic data, try as sensitivity analysis later
#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
# renv::status() to check 
# renv::snapshot() to add used and install to renv

@@

#standarsing behaviour variables
data <- data %>%
mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric()))

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
  filter(bd01 != 'no') %>% # EXCLUSION
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        )
  ) %>% 
  #Original paper used sum scores instead of FA. 
  mutate(
    ss_psycown = 
      across(starts_with("po"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_selfid = 
      across(starts_with("si"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_intention = 
      across(starts_with("it"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    #ss_donation as only measuring bd03
    ss_donation = 
      across(all_of(c("bd03")), ~as.numeric(.)) %>%
      rowSums(na.rm = TRUE)
    )

# To make bootstrapping run there cannot be any unordered factors
data.fit2 <- data.fit %>% 
  mutate(
    de01 = as.ordered(de01)
  ) %>% # and neither any factors at all
mutate(
  across(everything(),as.numeric)
)  # but this also good for plotting correlations

@@

@@

data.fit3 <- data.fit2 %>%
  select(-ss_psycown,
         -ss_selfid, 
         -ss_intention,
         -ss_donation)

corrplot(cor(data.fit3,use="pairwise.complete.obs"))
mcar_test(data.fit3 %>% select(-dttma))

@@

@@

labelsforplottingnull <- c(
  "ss_psycown" = "Psychological ownership",
  "ss_selfid"  = "Self-identity",
  "ss_intention" = "Intention",
  "ss_donation" = "Blood donation history"
)

lavaanPlot(
  model = fit.po02ss, 
   #labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
```

@@

```{r}
lavaanPlot(
  model = fit.po03ss, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))

@@

lavaanPlot(
  model = fit.po04ss, 
  labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))

@@

lavaanPlot(
  model = fit.po05ss, 
  labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))

@@

```{r}
labelsforplotting <- c(
  "psycown" = "Psychological ownership",
  "selfid"  = "Self-identity",
  "intention" = "Intention",
  "donation" = "Blood donation history"
)

lavaanPlot(
  model = fit.po02, 
  labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po03, 
  labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po04, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po05, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(worcs)
library(tidyverse)
library(gtsummary)
library(likert)
library(here)
library(corrplot)
library(lavaan)
library(lavaanPlot)
library(kableExtra)
library(naniar) # for missing data
library(patchwork)
library(gridExtra)
library(gt) # for table as plot
library(ggpmisc) 
library(ggcorrplot) #for corrplot
library(ggstats) # for likert plots

#library(semPlot) # update plots to semPlot later
#library(blavaan) # could not get it to work with this model and synthetic data, try as sensitivity analysis later
#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
# renv::status() to check 
# renv::snapshot() to add used and install to renv

@@

#standarsing behaviour variables
data.fit <- data %>%
mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric()))

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
  filter(bd01 != 'no') %>% # EXCLUSION
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        )
  ) %>% 
  #Original paper used sum scores instead of FA. 
  mutate(
    ss_psycown = 
      across(starts_with("po"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_selfid = 
      across(starts_with("si"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_intention = 
      across(starts_with("it"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_control = across(starts_with("ac"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_knowledge = across(starts_with("ai"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_selfinv = across(starts_with("iv"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    #ss_donation as only measuring bd03
    ss_donation = 
      across(all_of(c("bd03")), ~as.numeric(.)) %>%
      rowSums(na.rm = TRUE)
    )

# To make bootstrapping run there cannot be any unordered factors
data.fit2 <- data.fit %>% 
  #standarsing behaviour variables
  mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric())) %>% 
  #remove these as they are not used in modelling
  select(
    -ss_control,
    -ss_knowledge,
    ss_selfinv
    ) %>% 
  mutate(
    de01 = as.ordered(de01)
  ) %>% # and neither any factors at all
mutate(
  across(everything(),as.numeric)
)  # but this also good for plotting correlations

@@

## Demographics with correlation

tmp <- data.fit %>% 
  select(
        de01,
         de02,
         ss_control,
         ss_knowledge,
         ss_selfinv,
         bd03, 
         ss_psycown,
         ss_selfid,
         ss_intention 
         ) %>%
  select(ncol(.):1) %>% 
  mutate(
    across(everything(),as.numeric)
  ) %>% 
  rename(
        "Gender"=de01,
         "Age (years)" = de02,
         "Control" = ss_control,
         "Knowledge" =  ss_knowledge,
         "Self-investment" = ss_selfinv,
         "Past donations in 2 years" = bd03, 
         "Psychological ownership" = ss_psycown,
         "Self-identity" = ss_selfid,
         "Intention" = ss_intention 
           )

pt2 <- ggcorrplot(
  #compute correlatins from normalised values
  cor(tmp,use="pairwise.complete.obs"),
  show.diag = FALSE,
  outline.color = "white",
  p.mat = cor_pmat(tmp,use="pairwise.complete.obs"),
  sig.level=0.01,
  #insig = "",
  lab=TRUE,
  lab_size = 3,
  show.legend = TRUE,
  tl.cex = 10,
  digits = 1
 )

#compute mean and sd on original scale
tmp <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
  filter(bd01 != 'no') %>% # EXCLUSION
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        de01 = as.numeric(de01), # for counting correlation
       )
  ) %>% 
  #Original paper used sum scores instead of FA. 
  mutate(
    ss_psycown = 
      across(starts_with("po"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE),
    ss_selfid = 
      across(starts_with("si"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) ,
    ss_intention = 
      across(starts_with("it"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) ,
    ss_control = across(starts_with("ac"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) ,
    ss_knowledge = across(starts_with("ai"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE),
    ss_selfinv = across(starts_with("iv"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) ,
  ) %>% select(
    de01,de02,ss_control,ss_knowledge,ss_selfinv,bd03, ss_psycown,ss_selfid,ss_intention 
  ) %>% pivot_longer(cols = everything(), names_to = "Variable", values_to = "val") %>%
  group_by(Variable) %>% 
  summarize(mean = mean(val,na.rm=TRUE), sd = sd(val,na.rm=TRUE)
  ) %>% mutate(
    Mean = as.character(round(mean,2)),
    SD = as.character(round(sd,2))
  ) %>% select(-mean,-sd) %>%  
  mutate(
    Variable = factor(
      Variable,
      labels = c(
          "Gender",
         "Age (years)",
         "Control",
         "Knowledge",
         "Self-investment",
         "Past donations in 2 years", 
         "Psychological ownership",
         "Self-identity",
         "Intention" 
      ),
      levels = c(
        "de01",
        "de02",
        "ss_control",
                 "ss_knowledge",
                 "ss_selfinv",
                 "bd03", 
                 "ss_psycown",
                 "ss_selfid",
                 "ss_intention"),
      ordered = TRUE
    ),
    x=1,
  )  %>% arrange(Variable)

tmp[tmp$Variable == 'de01','Mean'] <- ' ' 
tmp[tmp$Variable == 'de01','SD'] <- ' '

#Plot tehm
tab1 <-  gt(tmp %>% select(-x)) %>% tab_options(table.font.size = 12,data_row.padding = px(6))
pt2 + wrap_table(tab1,space="free_y")
```

```{r}
ggsave(filename = here("figures/table1asFig.pdf"),units = "cm",width = 18,height = 10)
ggsave(filename = here("figures/table1asFig.png"),units = "cm",width = 18,height = 10)
```

@@

data.fit3 <- data.fit2 %>%
  select(-ss_psycown,
         -ss_selfid, 
         -ss_intention,
         -ss_donation,
         -ss_selfinv)

mcar_test(data.fit3 %>% select(-dttma))

@@

```{r}
corrplot(cor(data.fit3,use="pairwise.complete.obs"))
```

@@

labelsforplottingnull <- c(
  "ss_psycown" = "Psychological ownership",
  "ss_selfid"  = "Self-identity",
  "ss_intention" = "Intention",
  "ss_donation" = "Past donation behaviour"
)

lavaanPlot(
  model = fit.po02ss, 
   #labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
```

@@

```{r}
lavaanPlot(
  model = fit.po03ss, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))

@@

lavaanPlot(
  model = fit.po04ss, 
  labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))

@@

lavaanPlot(
  model = fit.po05ss, 
  labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))

@@

```{r}
labelsforplotting <- c(
  "psycown" = "Psychological ownership",
  "selfid"  = "Self-identity",
  "intention" = "Intention",
  "donation" = "Past donation behaviour"
)

lavaanPlot(
  model = fit.po02, 
  labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
   stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po03, 
  labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001,
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po04, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  stars = "regress"
  #sig=0.001,
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))

@@

```{r}
lavaanPlot(
  model = fit.po05, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  #sig=0.001
  stars = "regress"
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
